---
title: "Exploratory Report of Juvenile Snowshoe Hares in Bonanza Experimental Forest"
author: "Joe Walderman"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(broom)
library(effsize)
library(ggbeeswarm)
library(kableExtra)
```



### Introduction

The northern boreal forests are the habitat of the snowshow hare, a keystone species. This report utilizes a study on populations of the species in Bonanza Experimental Forest conducted by University of Alaska - Fairbanks from 1999 - 2012. This study utilized capture-recapture methods to develop an understanding of the population as a whole. Throughout the report, weight is compared based on sex of the snowshoe hairs. Further on, the relationship between weight and hind foot length is explored. 

### Data and Analysis

These relationships and comparisons are performed using summary statistics, data visualization, central tendency comparison, and linear regression modeling. This report utilized data collected from 1999-2012 of showshoe hare trappings in Bonanza Experimental Forest.



```{r}
hares <- read_csv(here("data", "bonanza_hares.csv"))


# Isolating hare trappings classified as juvenile and pulling the year of trapping for each observation from the date.

hares_dates <- hares %>%
  mutate(date_new = mdy(date)) %>%
  mutate(year_new = year(date_new))

hares_j <- hares_dates %>% 
  filter(age == "j")

hares_j_yearly <- hares_j %>% 
  group_by(year_new) %>% 
  summarize(count = n())

# Pulling vector for hare count

hares_count <- hares_j_yearly %>% 
  pull(count)

hares_j$year_new <- as.factor(hares_j$year_new)
```

### Exploratory Findings


```{r}
ggplot(data = hares_j_yearly, aes(x = year_new, y = count)) +
  geom_col() +
  labs(x = "Year",
       y = "Hares",
       title = "Annual Juvenile Hare Trappings in Bonanza Creek (1999-2012)",
       tag = "Figure 1: Number of juvenile showhow hares trapped each year in Bonanza Creek Experimental Forest from 1999-2012. Data: Kielland et al, 2017") +
  theme(plot.tag.position = "bottom",
        plot.tag = element_text(size = 5),
        plot.title = element_text(hjust = .5, size = 16))
```

**Figure 1.**: Number of juvenile showhow hares trapped each year in Bonanza Creek Experimental Forest from 1999-2012. Data: Kielland et al, 2017"


```{r}
#Getting summary statistics for yearly hare counts

mean_yearly <- mean(hares_count)
median_yearly <- median(hares_count)
min_yearly <- min(hares_count)
max_yearly <- max(hares_count)
```

The largest number of trappings (`r round(max_yearly,2)`) occurred in the first year of this study in 1999, and declined in subsequent years, though there are staggered upticks after a certain low amount is reached. Each rebound, though, is smaller every time it occurs. The lowest trapping count for a year in this study was `r round(min_yearly,2)`. It seems that there has been an overall decline in the hare population in these areas based on these samples, with intermittent periods of recovery followed by decline. Another point of interest in the notable difference between the mean `r round(mean_yearly,2)` and median `r round(median_yearly)` yearly hare trappings, with the median tending much lower.  Important to note in this data is that while there were hares trappings recorded in 2002 and 2009, none of these observations had a categorization for age, and thus are not included in this analysis. Further, to standardize this data accross years it is suggested that the same amount of days and traps be utilized for each year of collection.





### Comparing weights of male vs female juvenile hares at different capture sites


```{r}
#Expanding and clarifying variable labels

hares_j_sites <- hares_j %>% 
  mutate(grid_full = case_when(
    grid == "bonrip" ~ "Bonanza Riparian",
    grid == "bonmat" ~ "Bonanza Mature",
    grid == "bonbs" ~ "Bonanza Lowland"
  )) %>% 
  mutate(sex_full = case_when(
  sex == "f" ~ "Female",
  sex == "m" ~ "Male",
  sex == "NA" ~ "Unknown"
  ))
```


```{r}
ggplot(data = hares_j_sites, aes(x = sex_full, y = weight)) +
  geom_violin(aes(fill = sex_full), position = "dodge", alpha = 0.7) +
  geom_boxplot(fill = NA, width = .3, outlier.color = NA) +
  stat_summary(
    fun = mean,
    geom = "point",
    ) +
  labs(
    x = "Sex",
    y = "Weight (grams)"
  ) +
  theme_light() +
  facet_wrap(~grid_full)
  
```

**Figure 2.** Weight (g) observations for male and female juvenile hares, as well as those whose age has not been identified (NA), at three different collection sites. The width of the plots indicate the number of observations for a particular weight. Box endpoints indicate the 25th and 75th percentile values; the black line and black point withing the box indicate the median and mean value for each sex, respectively. Data: Kielland et al. 2017.





Males generally have a higher density of observations in higher weight recordings than females. The width of the plot represents the density of observations at that weight level. The boxplots indicate the sample sizes which can contribute to the variance and the mean which can more easily illustrate the differences between the two.





**Table 1.** Descriptive statistics (mean, standard deviation, and sample size) for male and female juvenile hares. Data: Kielland et al. 2017.

```{r descriptive_stats}
hares_j_table <- hares_j_sites %>%
  group_by(sex_full) %>% 
  filter(!is.na(weight)) %>% 
  filter(!is.na(sex)) %>%
  summarize(mean_weight = round(mean(weight, na.rm = TRUE),2),
            sd_weight = round(sd(weight, na.rm = TRUE),2),
            sample_size = n())

hares_j_table %>% 
  kable(col.names = c("Sex", "Mean Weight (g)", "Standard Deviation (g)", "Sample Size")) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE)
```




```{r, include=FALSE}
#Pulling vectors of weight for male and female juvenile hares
hares_female <- hares_j_sites %>% 
  filter(sex_full == "Female") %>% 
  pull(weight)

hares_male <- hares_j_sites %>% 
  filter(sex_full == "Male") %>% 
  pull(weight)

# Getting means and sds for male and female juvenile hares
male_mean <- mean(hares_male, na.rm = TRUE)
female_mean <- mean(hares_female, na.rm = TRUE)
male_sd <- sd(hares_male)
female_sd <- sd(hares_female)

#Plotting histograms and QQ plots to check assumptions of t-test
hist(hares_male) #does not look particularly normal
qqnorm(hares_male) #looks a bit curved
hist(hares_female) #looks fairly normal, if a bit skewed left
qqnorm(hares_female) #looks very linear

# T-test for male and female juvenile hares
hares_ttest <- t.test(hares_male, hares_female, na.rm = TRUE)

hares_ttest_tidy <- tidy(hares_ttest)

# Cohen's d effect size for male and female juvenile hares
hares_effsize <- cohen.d(hares_male, hares_female, na.rm = TRUE)
```

The absolute difference in the means is `r round(male_mean,2) - round(female_mean,2)` and the Welch's two sample t-test yeilds a p-value of t(`r round(hares_ttest_tidy$parameter,2)`) = `r round(hares_ttest_tidy$statistic,2)`, p = 0.007) which indicates there is a 0.7% chance that one might take sample two samples from the same population with means that are at least as different as the means observed here. This may suggest that the difference in means is significant. However,  along with the somewhat small Cohen's d Effect size `r round(hares_effsize$estimate,2)` does not corroborate with these results.The validity of these tests is based on assumptions of normality, however, and histograms and QQ plots for male penguins suggest that the distribution for that data may not be normal and satisfy these assumptions.




#### Relationships between juvenle weight and hind foot length

##### Linear Regression Model

```{r, include = FALSE}
hares_j_lm <- lm(weight ~ hindft, data = hares_j_sites)

hares_j_lm_tidy <- broom::tidy(hares_j_lm)

hares_j_lm_glance <- broom::glance(hares_j_lm)

hares_coef <- hares_j_lm_tidy$estimate[2]
hares_coef

summary(hares_j_lm) #returning the complete overview

plot(hares_j_lm) # *remember to include = false*

hares_j_cor <- cor.test(hares_j$hindft, hares_j$weight)

hares_j_cor_tidy <- tidy(hares_j_cor)
```

```{r}
ggplot(data = hares_j_sites, aes(x = hindft, y = weight)) +
  geom_point(size = 1.5) +  
  geom_smooth(method = "lm",
              color = "red",
              size = 0.5,
              fill = "gray10",
              alpha = 0.5) +
  labs(
    x = "Hind Foot Length (mm)",
    y = "Weight (g)",
    title = "Weight and Hind Foot Length in Juvenile Snowshow Hares",
    tag = "Figure 4: This graph explores the relationship between weight and hind food length in juvenile snowshoe hares captured in this study") +
      theme(plot.tag.position = "bottom",
        plot.tag = element_text(size = 6),
        plot.title = element_text(hjust = .5, size = 14))
```

**Figure 3.** Relationship between hind foot length (mm) and weight (g) for juvenile hairs. Points indicate individual hare observations. Linear model summary: ($\beta$ = `r round(hares_coef)`, F(`r hares_j_lm$df`_glance,`r hares_j_lm_glance$df.residual`) = `r round(hares_j_lm_glance$statistic,1)`, p < 0.001) with an R^2^ of `r round(hares_j_lm_glance$r.squared,3)`



I have some concerns here about the homoscedacisity of the relationship. It seems that the residuals are even up to about the 100mm hind foot length but beyond that the model seems to be heteroscedastic, as the distance of the residuals from the regression line increases.



- The hindfoot coefficient indicates we should expect hare weight (g) to increase by 9.5234 grams for every 1 mm increase in hind foot length

- The Multiple R^2^ value is 0.2988 - meaning that 29.9% of variance in weight is explained by hind foot length


Here, we see that there is a moderate positive correlation between snowshoe hare hindfoot length and weight(*r* = `r round(hares_j_cor$estimate,2)`, t(`r hares_j_cor$parameter`) = `r round(hares_j_cor$statistic,2)`, p < 0.001). 

I have some concerns here about the homoscedacisity of the relationship. It seems that the residuals are even up to about the 100mm hind foot length but beyond that the model seems to be heteroscedastic, as the distance of the residuals from the regression line increases.


### Summary

The following were suggested from the exploratory data analysis:

-Juvenile showshoe hares have fluctuated in population but overall declined from 1999 - 2012.

-Juvenile male snowshoe hares tend to weigh more than juvenile female snowshoe hares.

-There is a positive correlation between weight and hindd foot length with a moderate linear relationship.


### Citation

Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14 (Accessed 2020-11-21).