---
title: "a4task2jw"
author: "Joe Walderman"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
```


## Exploratory Findings: In this section I will be calculating juvenile hare trap counts for each year of the study.

### Step 1: Reading in the data

```{r}
hares <- read_csv(here("data", "bonanza_hares.csv"))
```

### Step 2:

```{r}
hares_dates <- hares %>%
  mutate(date_new = mdy(date)) %>%
  mutate(year_new = year(date_new))

hares_j <- hares_dates %>% 
  filter(age == "j")

hares_j_yearly <- hares_j %>% 
  group_by(year_new) %>% 
  summarize(count = n())

hares_j$year_new <- as.factor(hares_j$year_new)
```

### Step 3:

```{r}
ggplot(data = hares_j_yearly, aes(x = year_new, y = count)) +
  geom_col() +
  labs(x = "Year",
       y = "Hares",
       title = "Annual Juvenile Hare Trappings in Bonanza Creek (1999-2012)",
       tag = "Figure 1: Number of juvenile showhow hares trapped each year in Bonanza Creek Experimental Forest from 1999-2012. These numbers were obtained in a study conducted by University of Alaska - Fairbanks.") +
  theme(plot.tag.position = "bottom",
        plot.tag = element_text(size = 5),
        plot.title = element_text(hjust = .5, size = 16))
```

### Step 4:

```{r}
hares_j_sum <- hares_j_yearly %>% 
  summarize(mean_count = mean(count),
            median_count = median(count),
            minimum = min(count),
            maximum = max(count))

hares_j_sum
```

Some of the major takeaways from this graph is that the number of trappings declined after the first two years, though there are staggered upticks after a certain low amount is reached. Each rebound, though, is smaller every time it occurs.  Every tie thaThe lowest trapping count for a year in this study was `mean(hares_j_stats$count)` and


```{r}
hares_j_sites <- hares_j %>% 
  mutate(grid_full = case_when(
    grid == "bonrip" ~ "Bonanza Riparian",
    grid == "bonmat" ~ "Bonanza Mature",
    grid == "bonbs" ~ "Bonanza Lowland"
  )) %>% 
  mutate(sex_full = case_when(
  sex == "f" ~ "Female",
  sex == "m" ~ "Male",
  sex == "NA" ~ "Unknown"
  ))
```


```{r}
ggplot(data = hares_j_sites, aes(x = sex_full, y = weight, fill = sex_full)) +
  geom_violin(position = "dodge", alpha = 0.7) +
  geom_boxplot(width = .1) +
  labs(
    x = "Sex",
    y = "Weight (grams)"
  ) +
  theme_light() +
  facet_wrap(~grid_full)
  
```


```{r}
hares_j_stat <- hares_j_sites %>%
  group_by(sex_full) %>% 
  summarize(mean_weight = mean(weight, na.rm = TRUE),
            sd_weight = sd(weight, na.rm = TRUE),
            sample_size = n())

hares_j_stat
```


```{r}
hares_female <- hares_j_sites %>% 
  filter(sex_full == "Female") %>% 
  pull(weight)

hares_male <- hares_j_sites %>% 
  filter(sex_full == "Male") %>% 
  pull(weight)

t.test(hares_female, hares_male)
```

The actual difference in mean weight between female and male snowshoe hares is ``. 


